---
title: "Advanced visualization"
author: "Reka Toth "
ddate: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
 #workflowr::wflow_html:
    toc: false
    code_folding: "hide"
#output: workflowr::wflow_html:
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r message=FALSE, warning=FALSE}


if (!require(tidyr)) {
  install.packages("tidyr")
}
if (!require(dplyr)) {
  install.packages("dplyr")
}
if (!require(readr)) {
  install.packages("readr")
}
if (!require(ggplot2)) {
  install.packages("ggplot2")
}
if (!require(ggsci)) {
  install.packages("ggsci")
}
if (!require(pheatmap)) {
  install.packages("pheatmap")
}
if (!require(ggalluvial)) {
  install.packages("ggalluvial")
}
if (!require(cowplot)) {
  install.packages("cowplot")
}
if (!require(ggrepel)) {
  install.packages("ggrepel")
}
if (!require(ggpubr)) {
  install.packages("ggpubr")
}
if (!require(tibble)) {
  install.packages("tibble")
}

```

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(plotly)
library(ggplot2)
library(ggsci)
library(pheatmap)
library(ggalluvial)

```

## How to make publication ready figures?

1.  Selecting the type of visualization
2.  Making the plots visually appealing
    1.  ggplot2
    2.  Choosing and using palettes
    3.  supporting packages: ggpubr, ggrepl
    4.  Arrangement: patchwork, cowplot
3.  Special types of plots:
    1.  Heatmaps
    2.  Genomic plots
    3.  Alluvial plots
    4.  Upset plots

Data source: [CBioPortal](https://www.cbioportal.org/), [TCGA, LGG](https://www.cbioportal.org/study/summary?id=lgg_tcga_pan_can_atlas_2018)


```{r}
#| warning: false

LGG <- read.delim("data/lgg_tcga_pan_can_atlas_2018_clinical_data.tsv")

```

## Visualization types

#### One variable, continuous data

We would like to see, how the data is distributed in our study.

```{r}
#| warning: false

ggplot(LGG) + 
  geom_density(aes(Diagnosis.Age))+theme_bw()+ ggtitle("Density plot")+xlab("Age at diagnosis")

ggplot(LGG) + 
  geom_histogram(aes(Diagnosis.Age), fill="gold", color="black")+theme_bw()+ ggtitle("Histogram")+xlab("Age at diagnosis")

ggplot(LGG) + 
  geom_histogram(aes(Diagnosis.Age), bins = 10, fill="gold", color="black")+theme_bw()+ ggtitle("Histogram with diffierent bin size")+xlab("Age at diagnosis")



```

#### One variable, discrete data

We would like to see, how the data is distributed in our study.

```{r}
#| warning: false

ggplot(LGG) + 
  geom_bar(aes(Race.Category))+theme_bw()+ ggtitle("Bar plot")

LGG %>%
  filter(!is.na(Race.Category)) %>%
  mutate(categories="Race") %>%
  ggplot() +
  geom_bar(aes(x=categories,  fill=Race.Category)) + 
  coord_flip() + 
  theme_minimal() +
  theme(aspect.ratio=0.2, axis.title.y = element_blank())+ 
  ggtitle("Stacked bar plot")+
  scale_fill_futurama(name="Race category")

```

#### Two variables, continuous x and y

```{r}
#| warning: false

f <- ggplot(LGG, aes(x = Winter.Hypoxia.Score, y=Ragnum.Hypoxia.Score))

f + geom_point()+ggtitle("Geom_point")

f + geom_jitter()+ggtitle("Geom_jitter")



```

#### Two variables, discrete x and y

```{r}
#| warning: false

e <- LGG %>%
  filter(!is.na(Subtype)) %>%
 ggplot(aes(x = Subtype, y=MSI.MANTIS.Score))


e +
  geom_point()+ggtitle("Geom_point") +
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())

e +
  geom_jitter(width = 0.08)+ggtitle("Geom_jitter")+
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())


e +
  geom_boxplot()+ggtitle("Geom_boxplot")+
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())

e +
  geom_violin(fill="salmon")+ggtitle("Geom_violin")+
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())

e +
  geom_dotplot(binaxis = "y", stackdir =
"center", binwidth = 1/500)+ggtitle("Geom_dotplot")+
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())

e + geom_boxplot(width=0.4, fill="lightpink") + geom_jitter(width = 0.08, alpha=0.7, color="lightpink4") +
  theme_bw()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
    ylab("MSI Mantis Score")+
   theme(axis.title.x = element_blank())

```

#### discrete x , discrete y

```{r}

g <- LGG %>%
  filter(!is.na(Subtype) & !is.na(Race.Category)) %>%
  #count(Subtype, Race.Category) %>%
 ggplot(aes(x = Subtype, y=Race.Category)) 

g + geom_count(color="red4")+
   theme_minimal()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
   scale_fill_material("blue", name="Count")+
   theme(axis.title = element_blank(), axis.text.x = element_text(angle=45, hjust = 1))



 LGG %>%
    filter(!is.na(Subtype) & !is.na(Race.Category)) %>%
     count(Subtype, Race.Category) %>% 
   ggplot(aes(x = Subtype, y=Race.Category)) +
   geom_tile(aes(fill = n))+
   theme_minimal()+
   scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel")) + 
   scale_fill_material("blue", name="Count")+
   theme(axis.title = element_blank(), axis.text.x = element_text(angle=45, hjust = 1))


```

### Multiple variables

#### Multiple discrete variables with the same categories

For example the change of categorical variables in multiple years. Use instead of pie chart.

```{r}

LGG %>%
  filter(!is.na(Subtype)) %>%
ggplot() +
geom_bar(aes(x=Subtype,  fill=Cancer.Type.Detailed), width = 0.5, position = "fill", color="black") + coord_flip()  + theme_bw() + scale_fill_futurama(name="Cancer type")+ylab("Proportion")+theme(axis.title.y = element_blank(), panel.border = element_blank(), aspect.ratio = 0.8)+scale_x_discrete(labels=c("LGG_IDHwt" = "IDH WT", "LGG_IDHmut-non-codel" = "IDH mutant, not 1p/19q codel", "LGG_IDHmut-codel" = "IDH mutant,  1p/19q codel"))

```

#### Multiple continuous variables

Example: gene expression values

```{r, eval=FALSE}


library(TCGAbiolinks)
library(SummarizedExperiment)


query <- GDCquery(
    project = "TCGA-LGG",
    data.category = "Transcriptome Profiling",
    data.type = "Gene Expression Quantification", 
    workflow.type = "STAR - Counts"
)
GDCdownload(query = query)
data <- GDCprepare(query = query)
saveRDS(data, "data/LGG_GE.RDS")

ge1 <- data[order(rowSds(assay(data, "fpkm_uq_unstrand")), decreasing = T)[1:500],]
ge1 <- assay(ge1, "fpkm_uq_unstrand")
ge1 <- ge1+0.001
ge1 <- log2(ge1)
saveRDS(ge1, "data/LGG_GE_small_mat.RDS")


```

```{r}

ge <- readRDS("data/LGG_GE_small_mat.RDS")


pheatmap(ge[1:50,], scale = "row", show_rownames = F, show_colnames = F, colorRampPalette(c("navy", "white", "red"))(50))

#pheatmap(ge[1:50,], scale = "row", show_rownames = F, show_colnames = F, colorRampPalette(c("navy", "white", "red"))(50), annotation_col = LGG_anno[,c("Sex", "Cancer.Type.Detailed")])


```


### Special types

#### Upset plot

Use UpSet plot instead of Venn Diagram!

```{r}

library(UpSetR)

mutation_data <- matrix(sample(c(0,1), 10000, replace = T, prob = c(0.8, 0.2)), nrow=1000)
colnames(mutation_data) <- c("TP53", "KMT2C", "KMT2D", "ARID1A", "PIK3CA", "PTEN", "EGFR", "ERBB2", "KRAS",  "BRAF")
mutation_data <- as.data.frame(mutation_data)

upset(mutation_data, nsets = 10, mb.ratio = c(0.55, 0.45), order.by = "freq")


```

#### Alluvial plot

To visualize change over time.

```{r}
data(majors)
#> pivot_wider(majors, names_from = semester, values_from = curriculum)

majors$curriculum <- as.factor(majors$curriculum)
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum, label = curriculum)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray") +
  geom_stratum() +
  theme(legend.position = "bottom") +
  ggtitle("student curricula across several semesters")


```

```{r}

clinical_alluvial <- readRDS("data/clinical_alluvial.RDS")


g <- ggplot(clinical_alluvial ,
                aes(x = State, stratum = cluster, alluvium = case_barcode,
                    fill = cluster , label = cluster )) +
  scale_fill_igv()+
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray") +
  geom_stratum() +
  theme(legend.position = "bottom") +
  ggtitle("Cluster consistency of IDH wt GBM recurrency")



```

```{r}

ggplot(as.data.frame(Titanic),
       aes(y = Freq,
           axis1 = Class, axis2 = Sex, axis3 = Age, axis4 = Survived)) +
  geom_stratum(width = 1) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Class", "Sex", "Age", "Survived"))

ggplot(as.data.frame(Titanic),
       aes(y = Freq,
          axis1 = Class, axis2 = Sex, axis3 = Age, axis4 = Survived)) +
  geom_flow(aes(fill = Survived)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Class", "Sex", "Age", "Survived")) 

ggplot(as.data.frame(Titanic),
       aes(y = Freq,
          axis1= Sex, axis2 = Age, axis3 = Survived)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", aes(fill = Survived)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c( "Sex", "Age", "Survived")) 

```

