---
title: "ggplot exercises"
author: "Reka Toth "
ddate: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: false
    code_folding: "hide"
#output: workflowr::wflow_html:
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


# Exercises

1. Use the iris dataset.
2. plot a scatterplot using sepal.length and sepal.width
3. Color the dots by Species
4. Add linear regression curve for the whole dataset
5. Add linear regression curve colored by species
6. plot boxplots for sepal length, one for each species, with different color
7. plot dots on the boxplot
8. Change the colors to red, blue and green.
9. change axis labels
10 Flip the coordinates
11. Change the theme to theme_bw

```{r}
library(ggplot2)
data(iris)
iris
ggplot(iris, aes(Sepal.Length, Sepal.Width))+geom_point(pch=21, color="black", fill=alpha("red", 0.2), size=4)+stat_smooth(method="lm")+scale_x_continuous(name = "Sepal Width")+theme(axis.title.y = element_blank())




```
